---
output:
  pdf_document: default
  html_document: default
  word_document: default
---

STAT430 Homework #11: Due Saturday, May 11, 2019. 
========================================================

#### Name: Oliver Shanklin

*********************************************************************************
\newcommand{\prob}{\overset{P}{\to}}

0. We finish **Chapter 10** on hypothesis testing.     


1. Complete **Exercise 10.23** of the text. For part (c), give a complete **RANTDRC** test of the appropriate hypothesis.  

*********************************************************************************
  
**Answer:**

#### a)

$H_a : \mu_1 \ne \mu_2$

$H_o : \mu_1 = \mu_2$

#### b)

This would be a two tailed test because this a difference.

### c)

**R**: Do the ranges of deer populations differ?

**A**: $H_a : \mu_1 \ne \mu_2$

**N**: $H_o : \mu_1 = \mu_2$

**T**: $Z = \frac{\mu_1 - \mu_2}{\sqrt{\frac{S_1^2}{n_1}+\frac{S_2^2}{n_2}}} = -0.95358$

**D**: $Z \sim N(0,1)$ where $\pm Z_{\alpha/2} = \pm1.645$

**R**: $Z = -0.95358 > -1.645$

**C**: We sould Fail to Reject $H_o$, so the deer do not travel significantly different differences.


*********************************************************************************

2. Complete **Exercise 10.24** of the text, giving a complete **RANTDRC** test of the appropriate hypothesis.  


*********************************************************************************
  
**Answer:**

**R**: Is the percentage reported too high?

**A**: $H_a : p < 0.15$

**N**: $H_o : p = 0.15$

**T**: $Z = \frac{\hat p - p}{\sqrt{\frac{\hat p(1- \hat p)}{n}}} = \frac{0.13-0.15}{\sqrt{\frac{.13(.87)}{100}}} = -0.5947$

**D**: $Z\sim N(0,1)$ where $Z_{\alpha/2} = -1.645$

**R**: $-0.5947 > -1.645$

**C**: We would Fail to Reject $H_o$.


*********************************************************************************


3. Complete **Exercise 10.33** of the text, giving a complete **RANTDRC** test of the appropriate hypothesis.  


*********************************************************************************
  
**Answer:**

**R**: Republicans support it more. 

**A**: $H_a: p_1 > p_2$

**N**: $H_o: p_1 = p_2$

**T**: $Z = \frac{\hat p_1 - \hat p_2}{\sqrt{\frac{\hat p_1(1-\hat p_1)}{n}+ \frac{\hat p_2(1-\hat p_2)}{n}}} = 1.504$

**D**: $Z \sim N(0,1)$ where $Z_\alpha = 1.645$

**R**: $1.504 < 1.645$

**C**: So we would Fail to Reject $H_o$.
   
*********************************************************************************

4. Complete **Exercise 10.38** of the text. 

*********************************************************************************
  
**Answer:**

$\beta = P(T \notin RR, H_o is false)$

$P\left(\frac{\bar X - \mu_o}{s/\sqrt{n}} \ge -2.33\right)$
$= P\left({\bar X } \ge \mu_o - 2.33(s/\sqrt{n})\right)$
$= P\left({\bar X } \ge 61.3752\right)$
$= P\left(\frac{\bar X - 60}{\sqrt{S^2/n}} \ge \frac{61.37 - 60}{\sqrt{S^2/n}}\right)$
$= P\left({Z } \ge 1.21\right) = 0.1132$


*********************************************************************************

5. Complete **Exercise 10.45** of the text. 

*********************************************************************************
  
**Answer:**

$\left[(\mu_1 - \mu_2) \pm Z_{0.005}\sqrt{\frac{S_1^2}{n_1}+\frac{S_2^2}{n_2}} \right]$

#### a)

After computing the interval, $0$ is not in the interval.

#### b)

So $H_o$ would be rejected because the target is $\mu_1 - \mu_2 =0$.

#### c)

We have the same conclusion.
*********************************************************************************

6. Complete **Exercise 10.50** of the text.  

*********************************************************************************
  
**Answer:**

$\hat p = 0.58$, $sd = .11$, $p=0.6$

$H_a: p < 0.6$

$H_o: p = 0.6$

$T=\frac{0.58 -0.6}{\sqrt{0.11^2/120}} = -1.9917 \sim N(0,1)$

```{r}
pnorm(-1.9917)
```

So we would reject $H_o$ at $\alpha = 0.1$ and say it is unprofitable.

*********************************************************************************


7. Complete **Exercise 10.54** of the text.  


*********************************************************************************
  
**Answer:**

$p = 0.85$, $\hat p = 0.96$, $n = 300$, $\alpha = 0.01$

$$Z = \frac{\hat p - p}{\sqrt{\frac{\hat p}{1 - \hat p}/n}} \sim N(0,1)$$

```{r}
pnorm(9.7227)
```

Since the p-value is so small, we would reject $H_o$ in almost every situtaion. So the reported proportion is too low.

*********************************************************************************
   

8. Complete **Exercise 10.106** of the text. 


*********************************************************************************

**Answer:**

$\hat \theta_1 = 76/200$

$\hat \theta_2 = 53/200$

$\hat \theta_3 = 59/200$

$\hat \theta_4 = 48/200$

$\hat \theta_0 = 236/800$

$H_o: \theta_1 = \theta_2 = \theta_3 = \theta_4 = \theta_0$

$H_a: \theta_1, \theta_2, \theta_3, \theta_4$, at least one is different.


$T = -2ln(LR) \sim \chi^2_3$


Max Likelihood under $H_0$ =  $\prod_{i=1}^4{ 200 \choose y_i} \theta_0^{y_i} (1-\theta_0)^{200-y_i}$

```{r}
ML1 <- dbinom(76,200,236/800)*dbinom(53,200,236/800)*dbinom(59,200,236/800)*dbinom(48,200,236/800)
```

Max Likelihood = $\prod_{i=1}^4{ 200 \choose y_i} \theta_i^{y_i} (1-\theta_i)^{200-y_i}$

```{r}
ML2 <- dbinom(76,200,76/200)*dbinom(53,200,53/200)*dbinom(59,200,59/200)*dbinom(48,200,48/200)
```

```{r}
Tstat <- -2*log(ML1/ML2)

Tstat

(qchisq(0.95, 3)<Tstat)
```

Since $T$ is larger than $\alpha = 0.05$ for a $\chi^2_3$ we reject $H_0$ and say that at least one ward votes differently.


*********************************************************************************
   


**Optional problems:**   We have touched on nearly all of Chapter 10, especially sections 10.2, 10.3, 10.4, 10.5, 10.6, and 10.11.  Any of the (non-applet) exercises in those sections are good to review.  Less emphasis on the exact tests in 10.8 and 10.9, though these are just applications of the normal-theory distributions that we studied in the first part of the course and applied in the confidence interval section.  

